---
output: github_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  fig.path = "./man/figures/README-",
  fig.align = "center",
  collapse = TRUE,
  comment = "#>"
)
```

# aisstream

<!-- badges: start -->
<!-- badges: end -->

A thin R client for the [AISStream.io](https://aisstream.io) live vessel-tracking
(AIS) WebSocket feed, built on
[connectcore](https://github.com/dereckscompany/connectcore). Open one socket,
subscribe with bounding boxes plus optional MMSI / message-type filters, and handle
raw frames Node-ws style. Reconnect, re-subscribe, keepalive and a silence watchdog
are inherited; parse helpers and a durable NDJSON recorder are included.

## Why it is shaped this way

AISStream is **WebSocket-only** (no REST). On connect you must send one JSON
subscription within three seconds or the server drops you, and — the rule that
shapes everything — it **monitors your TCP read queue and closes the connection if
it backs up**. The whole-world feed runs at roughly 300 messages a second, so the
message hot path cannot afford a per-message JSON parse.

So `aisstream` keeps the data path **parse-free**: `.dispatch()` emits each frame as
the raw string and only does a cheap prefix check to split off error frames. You
record the raw frames now (see [`ndjson_sink()`](#recording)) and parse them
offline. The connection machinery — the 3-second subscribe, reconnect-and-
resubscribe, keepalive, the silence watchdog — all comes from `connectcore`.

## Installation

This package uses [renv](https://rstudio.github.io/renv/):

```r
renv::install("dereckscompany/aisstream")
```

## Constructing a client and inspecting the subscription

Everything below is **network-free** — no socket is opened. We build a client and
look at the exact subscription frame it would send.

```{r setup}
library(aisstream)

ais <- AisStream$new(
  api_key = "DEMO-KEY", # normally read from the AISSTREAM_API_KEY env var
  bounding_boxes = list(
    # a box is two opposite [lat, lon] corners; pass it JS-object style ...
    list(min_lat = 40.4, min_lon = -74.3, max_lat = 41.0, max_lon = -73.7),
    # ... or as raw nested corners (order does not matter).
    list(c(50.0, -1.5), c(51.0, 1.5))
  ),
  message_types = "PositionReport",
  ship_mmsi = c("366998510", "367123450")
)

# The exact JSON sent on (re)connect — APIKey + BoundingBoxes always, filters only
# when non-empty (an empty filter is omitted, never sent as []).
cat(ais$subscription_frame())
```

Filters are omitted entirely when empty:

```{r minimal}
minimal <- AisStream$new(
  api_key = "DEMO-KEY",
  bounding_boxes = list(list(min_lat = -90, min_lon = -180, max_lat = 90, max_lon = 180))
)
cat(minimal$subscription_frame())
```

## Handling frames

Register handlers Node-ws style, then `$run()` to pump the event loop. `"message"`
carries the **raw** string (parse it yourself if you must); `"error"` carries an
AISStream error frame.

```{r handlers, eval=FALSE}
ais$on("message", function(raw) {
  pr <- as_position_report(parse_ais(raw))
  print(pr[, .(mmsi, ship_name, latitude, longitude, sog, cog)])
})
ais$on("error", function(raw) message("AIS error: ", raw))

ais$run() # blocks, pumping the loop; Ctrl-C to stop
```

## Parsing frames offline

`parse_ais()` is JSON.parse for AIS, and the flatteners pull the stable common
fields into a `data.table`. The `time_utc` field arrives in Go's (non-ISO) format
and is parsed specially.

```{r parse}
frame <- paste0(
  '{"MessageType":"PositionReport",',
  '"MetaData":{"MMSI":368207620,"ShipName":"OCEAN TITAN  ",',
  '"latitude":40.12,"longitude":-74.21,',
  '"time_utc":"2022-12-29 18:22:32.318353 +0000 UTC"},',
  '"Message":{"PositionReport":{"Sog":12.3,"Cog":89.1,',
  '"TrueHeading":90,"NavigationalStatus":0,"RateOfTurn":-2}}}'
)

as_position_report(parse_ais(frame))
```

## Recording {#recording}

The proven "read fast or get dropped" pattern: `ndjson_sink()` returns a `"message"`
handler you register Node-ws style, then run. It appends each raw frame to an hourly
NDJSON file (rolling by UTC hour, flushing every few seconds) doing the minimum on
the hot path.

```{r record, eval=FALSE}
ais <- AisStream$new(
  bounding_boxes = list(list(min_lat = -90, min_lon = -180, max_lat = 90, max_lon = 180)),
  message_types = "PositionReport"
)
ais$on("message", ndjson_sink(dir = "ais-data"))
ais$run()
```

See `vignette("recording-ais")` for the full rationale.

## The message types

```{r types}
length(AIS_MESSAGE_TYPES)
head(unlist(AIS_MESSAGE_TYPES, use.names = FALSE))
```

## License

MIT
