---
title: "Recording AIS"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Recording AIS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(aisstream)
```

## Read fast, or get dropped

AISStream.io delivers AIS (vessel-tracking) messages over a single WebSocket. Two
properties of the feed dictate how a client must behave:

1. **You must subscribe within three seconds** of connecting, or the server closes
   the socket. The subscription is one JSON frame naming your bounding boxes (and
   optional filters).
2. **The server watches your TCP read queue and closes the connection if it backs
   up.** The whole-world feed runs at roughly 300 messages a second. If your handler
   does real work per message — parsing JSON, building a `data.table`, touching a
   database — the read queue grows and you get dropped.

`aisstream` answers both. The first is handled for you: `AisStream` re-sends the
subscription on every (re)connect via `connectcore`'s `.resubscribe()` seam, so the
3-second deadline is always met, including after a reconnect. The second shapes the
whole design — **the message hot path never parses**. `.dispatch()` emits each frame
as the raw string and does only a cheap prefix check to split off error frames.

The practical consequence: don't parse in the live handler. **Record the raw frames
now, parse them offline.** That is exactly what `record_to_ndjson()` does.

## The recorder

`record_to_ndjson()` wires a durable recorder onto a client and returns the client so
you can chain `$run()`:

```{r recorder, eval = FALSE}
ais <- AisStream$new(
  # AISSTREAM_API_KEY is read from the environment by default
  bounding_boxes = list(list(min_lat = -90, min_lon = -180, max_lat = 90, max_lon = 180)),
  message_types = "PositionReport"
)

record_to_ndjson(ais, dir = "ais-data", flush_seconds = 5)
ais$run() # blocks, pumping the event loop; Ctrl-C to stop
```

What it does, and why each part matters:

- **Opens the first hourly file before any message arrives**, so there is never a
  window where a frame has nowhere to go.
- **Appends each raw frame** with `connectcore::ws_file_sink()` — a one-line write,
  the minimum work, so the socket drains fast enough that the server never drops you.
  One JSON frame per line (NDJSON).
- **Flushes and rolls on the `later` loop**, not on the hot path. Every
  `flush_seconds` it flushes the buffer to disk; on the turn of a UTC hour it closes
  the current file and opens the next (`ais-YYYY-MM-DDTHH.ndjson`). A crash costs at
  most the last unflushed window.

Files are named by UTC hour, and `ndjson_hour_path()` lets you predict the path:

```{r hour-path}
ndjson_hour_path("ais-data", at = lubridate::ymd_hms("2024-03-15 09:42:11", tz = "UTC"))
```

## Parsing the recording offline

Once recorded, parse at your leisure — this is where the JSON work belongs. Read a
file line by line and turn each frame into a tidy row:

```{r parse}
# A line as it would appear in a recorded NDJSON file.
line <- paste0(
  '{"MessageType":"PositionReport",',
  '"MetaData":{"MMSI":368207620,"ShipName":"OCEAN TITAN  ",',
  '"latitude":40.12,"longitude":-74.21,',
  '"time_utc":"2022-12-29 18:22:32.318353 +0000 UTC"},',
  '"Message":{"PositionReport":{"Sog":12.3,"Cog":89.1,',
  '"TrueHeading":90,"NavigationalStatus":0,"RateOfTurn":-2}}}'
)

parsed <- parse_ais(line)
as_position_report(parsed)
```

`parse_ais()` is deliberately generic (JSON.parse for AIS): it does not model the 25
message bodies, because the AISStream API is in beta and its bodies shift. The
flatteners (`ais_metadata()`, `as_position_report()`) model only the stable common
fields, and `parse_go_time()` handles the Go-format `time_utc` (which is **not** ISO
8601) — note the sub-second precision is preserved:

```{r go-time}
parse_go_time("2022-12-29 18:22:32.318353 +0000 UTC")
```

A whole file folds into one table by reading its lines and row-binding:

```{r fold, eval = FALSE}
lines <- readLines("ais-data/ais-2024-03-15T09.ndjson")
positions <- data.table::rbindlist(
  lapply(lines, function(l) as_position_report(parse_ais(l))),
  fill = TRUE
)
```

## Changing the subscription live

A re-sent subscription is a full **swap-and-replace** (not a merge), rate-limited to
about once a second. `update_subscription()` replaces the stored parameters and, if
the socket is open, re-sends immediately:

```{r update, eval = FALSE}
ais$update_subscription(
  bounding_boxes = list(list(min_lat = 50, min_lon = -1.5, max_lat = 51, max_lon = 1.5))
)
```

That is the whole pattern: subscribe with bounding boxes, record raw frames fast,
parse offline. The connection lifecycle takes care of itself.
